---
title: "Preparing Data for sube"
vignette: >
  %\VignetteIndexEntry{Preparing Data for sube}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

`sube` works best when the data contracts are explicit. In practice, most
projects need four input families:

1. Supply-use table data in long form or workbook form.
2. Product and industry correspondence tables.
3. Industry-level input metrics used for multiplier and elasticity
   computation.
4. A prepared modeling table for SUBE regressions.

Whether your data comes from WIOD, FIGARO, or another national statistical
office, the same canonical format applies. The package ships importers for
WIOD workbooks (`import_suts()`) and FIGARO CSV flatfiles (`read_figaro()`),
but any supply-use data reshaped into the canonical long format described
below will flow through the same downstream pipeline.

The package deliberately separates matrix construction from econometric
estimation. That keeps the Leontief benchmark workflow usable even when the
modeling table is assembled elsewhere.

The shipped example objects map directly to those four families:

- `sube_example_data("sut_data")` -> normalized supply-use table input
- `sube_example_data("cpa_map")` and `sube_example_data("ind_map")` -> product and industry correspondence tables
- `sube_example_data("inputs")` -> compute-layer input metrics
- `sube_example_data("model_data")` -> prepared modeling table for `estimate_elasticities()`

That sample-data path is the easiest way to understand the required contracts before switching to external research data.

## Canonical SUT Format

All SUT data must resolve to a long-format data frame with exactly 7 columns. Each row represents one cell of the supply-use table. Supply and use tables are stacked in the same data frame and distinguished by the TYPE column.

| Column | Type | Semantics | Example |
|--------|------|-----------|---------|
| REP | character | Reporting country - the economy whose supply-use table is being described | `"AAA"` |
| PAR | character | Partner country - the economy supplying or demanding the product. Equal to REP for the domestic block | `"AAA"` |
| CPA | character | Product code - identifies the good or service (CPA/NACE product classification) | `"P1"` |
| VAR | character | Industry or final-demand variable - identifies the column in the SUT (an industry code or a final-demand code) | `"I1"`, `"FU_bas"` |
| VALUE | numeric | Cell value in the supply-use table, in the unit of the source data (typically millions of current-price currency units) | `10` |
| YEAR | integer | Reference year of the data | `2020` |
| TYPE | character | Table type: `"SUP"` for supply table, `"USE"` for use table | `"SUP"` |

The shipped example shows this structure concretely:

```{r}
sube_example_data("sut_data")
```

### Column name synonyms

Column names in the 7-column SUT table are matched after uppercasing only - there is no synonym resolution for the SUT columns themselves. Synonyms apply exclusively to **mapping table columns** (`cpa_map`, `ind_map`) passed to functions such as `import_suts()`. All matching is case-insensitive because column names are uppercased on import.

| Canonical Name | Accepted Synonyms |
|----------------|-------------------|
| CPA | CPA, CPA56, CPA_CODE |
| CPAAGG | CPAAGG, CPA_AGG, PRODUCT, PRODUCT_AGG |
| VARS | VARS, VAR, INDUSTRY, IND, CODE, NACE, NACE_R2 |
| INDAGG | INDAGG, IND_AGG, INDUSTRY_AGG, SECTOR |

## Satellite Vector Inputs

The `inputs` table is researcher-supplied from national accounts or external sources - the package does **not** derive these values from the SUT data. `compute_sube()` requires it as a separate argument alongside the matrix bundle.

| Column | Type | What It Measures | Required? | Source |
|--------|------|-----------------|-----------|--------|
| YEAR | integer | Reference year | Yes | Must match SUT YEAR values |
| REP | character | Country code | Yes | Must match SUT REP values |
| INDUSTRY (or synonym) | character | Industry identifier | Yes | Must match VAR codes in the SUT |
| GO | numeric | Gross output - total production value by industry | Yes | Researcher-supplied (e.g. national accounts) |
| VA | numeric | Value added - GDP contribution by industry | Optional | Researcher-supplied |
| EMP | numeric | Employment - workers or hours by industry | Optional | Researcher-supplied |
| CO2 | numeric | CO2 emissions by industry | Optional | Researcher-supplied |

The industry identifier column accepts synonyms: IND, INDUSTRY, INDUSTRIES, INDAGG (checked in that order; first match wins). GO is the only metric always required. VA, EMP, and CO2 are computed when present in the data frame.

```{r}
sube_example_data("inputs")
```

## Bring Your Own Data

This section walks through reshaping arbitrary supply-use data into the canonical format expected by the package.

### SUT table preparation

1. Identify which columns in your source data correspond to REP, PAR, CPA, YEAR, and TYPE.
2. Rename or map columns to the canonical names, or supply them with any casing - `import_suts()` will uppercase all column names on import.
3. Choose a layout: long format (all 7 columns including VAR and VALUE) or wide format (5 identity columns with industry codes as column headers). `import_suts()` accepts both.
4. If using wide format: be aware that aggregate demand columns (e.g. total final use) will be melted as industry columns unless they match known WIOD aggregates. Remove or rename them before import if needed.
5. Call `import_suts()` on the file path. Inspect the result with `head()` and `str()` to verify all 7 columns are present and values look correct.

### Satellite vector preparation

1. Obtain gross output (GO) by industry from national accounts or your preferred source. This column is required.
2. Optionally gather value added (VA), employment (EMP), and CO2 emissions (CO2) by industry.
3. Structure the data with columns YEAR, REP, and one industry identifier column (any of: IND, INDUSTRY, INDUSTRIES, or INDAGG).
4. Ensure YEAR and REP values match those in your SUT data exactly.
5. Provide one row per (YEAR, REP, INDUSTRY) combination. Pass the resulting data frame to the `inputs` argument of `compute_sube()`.

The remaining sections walk through the shipped example objects in detail, illustrating each data family in the preparation sequence.

## Supply-use data

The `sut_data` example below shows the 7-column canonical format described
above in practice - the shipped object already conforms to the contract, so
you can inspect its shape before reshaping your own source.

Imported SUT data should resolve to these columns:

```{r}
names(sube_example_data("sut_data"))
```

Use `import_suts()` for normalized CSV files or WIOD-style workbooks, then
filter with `extract_domestic_block()` when the domestic block is required.

For large comparative projects, it is worth standardizing country, partner,
year, variable, and product codes before import. The package will normalize
column names, but it does not guess missing structural fields.

## Mapping tables

Mapping tables bridge the canonical SUT columns to the aggregated matrix
dimensions needed by `build_matrices()`. The synonym resolution described
above applies here - use any of the accepted column-name variants for the
source code and aggregated code.

The product and industry maps can use simple two-column tables:

```{r}
sube_example_data("cpa_map")
sube_example_data("ind_map")
```

The package normalizes names internally, but the first two columns should still
represent source code and aggregated code.

In practice, these mapping tables determine the resolution of the full
workflow. If the package-level products differ from the paper-level products,
the comparison plots will reflect that aggregation choice.

## Input metrics

The `inputs` example provides the satellite vectors described in the Satellite
Vector Inputs section above - researcher-supplied series (GO required; VA,
EMP, CO2 optional) joined to the matrix bundle during compute.

The computation stage needs `YEAR`, `REP`, one industry identifier column, and
at least `GO`. Additional metrics such as `VA`, `EMP`, and `CO2` are computed
when present.

```{r}
sube_example_data("inputs")
```

## Modeling table

The modeling table combines matrix outputs with satellite-derived multipliers
for econometric estimation - it is the last-mile object assembled after the
Leontief benchmark layer is stable.

`estimate_elasticities()` uses a prepared table with:

- country and year identifiers
- one entity column, such as industries or products
- response variables such as `GO`, `VA`, `EMP`, or `CO2`
- predictor columns for the supply-use based explanatory structure

```{r}
names(sube_example_data("model_data"))
```

This design is intentional. Some projects will derive the modeling table from
the same supply-use source, while others will merge additional covariates or
paper-specific filters before estimation.

## Recommended preparation strategy

For reusable workflows, it is usually best to prepare data in this order. The shipped examples follow the same sequence, so you can inspect each object family before substituting your own files:

1. standardize raw supply-use inputs,
2. verify correspondence tables,
3. build the domestic matrix bundle,
4. compute Leontief benchmark results, and
5. construct the modeling table only after the benchmark layer is stable.
