---
title: "Preparing Data for sube"
vignette: >
  %\VignetteIndexEntry{Preparing Data for sube}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

`sube` expects three input families:

1. Supply-use table data in long form or workbook form.
2. Product and industry correspondence tables.
3. Industry-level input metrics used for multiplier and elasticity
   computation.

## Supply-use data

Imported SUT data should resolve to these columns:

```{r}
names(sube_example_data("sut_data"))
```

Use `import_suts()` for normalized CSV files or WIOD-style workbooks, then
filter with `extract_domestic_block()` when the domestic block is required.

## Mapping tables

The product and industry maps can use simple two-column tables:

```{r}
sube_example_data("cpa_map")
sube_example_data("ind_map")
```

The package normalizes names internally, but the first two columns should still
represent source code and aggregated code.

## Input metrics

The computation stage needs `YEAR`, `REP`, one industry identifier column, and
at least `GO`. Additional metrics such as `VA`, `EMP`, and `CO2` are computed
when present.

```{r}
sube_example_data("inputs")
```
