---
title: "Preparing Data for sube"
vignette: >
  %\VignetteIndexEntry{Preparing Data for sube}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

`sube` works best when the data contracts are explicit. In practice, most
projects need four input families:

1. Supply-use table data in long form or workbook form.
2. Product and industry correspondence tables.
3. Industry-level input metrics used for multiplier and elasticity
   computation.
4. A prepared modeling table for SUBE regressions.

The package deliberately separates matrix construction from econometric
estimation. That keeps the Leontief benchmark workflow usable even when the
modeling table is assembled elsewhere.

The shipped example objects map directly to those four families:

- `sube_example_data("sut_data")` -> normalized supply-use table input
- `sube_example_data("cpa_map")` and `sube_example_data("ind_map")` -> product and industry correspondence tables
- `sube_example_data("inputs")` -> compute-layer input metrics
- `sube_example_data("model_data")` -> prepared modeling table for `estimate_elasticities()`

That sample-data path is the easiest way to understand the required contracts before switching to external research data.

## Supply-use data

Imported SUT data should resolve to these columns:

```{r}
names(sube_example_data("sut_data"))
```

Use `import_suts()` for normalized CSV files or WIOD-style workbooks, then
filter with `extract_domestic_block()` when the domestic block is required.

For large comparative projects, it is worth standardizing country, partner,
year, variable, and product codes before import. The package will normalize
column names, but it does not guess missing structural fields.

## Mapping tables

The product and industry maps can use simple two-column tables:

```{r}
sube_example_data("cpa_map")
sube_example_data("ind_map")
```

The package normalizes names internally, but the first two columns should still
represent source code and aggregated code.

In practice, these mapping tables determine the resolution of the full
workflow. If the package-level products differ from the paper-level products,
the comparison plots will reflect that aggregation choice.

## Input metrics

The computation stage needs `YEAR`, `REP`, one industry identifier column, and
at least `GO`. Additional metrics such as `VA`, `EMP`, and `CO2` are computed
when present.

```{r}
sube_example_data("inputs")
```

## Modeling table

`estimate_elasticities()` uses a prepared table with:

- country and year identifiers
- one entity column, such as industries or products
- response variables such as `GO`, `VA`, `EMP`, or `CO2`
- predictor columns for the supply-use based explanatory structure

```{r}
names(sube_example_data("model_data"))
```

This design is intentional. Some projects will derive the modeling table from
the same supply-use source, while others will merge additional covariates or
paper-specific filters before estimation.

## Recommended preparation strategy

For reusable workflows, it is usually best to prepare data in this order. The shipped examples follow the same sequence, so you can inspect each object family before substituting your own files:

1. standardize raw supply-use inputs,
2. verify correspondence tables,
3. build the domestic matrix bundle,
4. compute Leontief benchmark results, and
5. construct the modeling table only after the benchmark layer is stable.
