---
title: "FIGARO End-to-End Workflow"
vignette: >
  %\VignetteIndexEntry{FIGARO End-to-End Workflow}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  eval = FALSE
)
library(sube)
```

# 1. What FIGARO is and what this vignette covers

FIGARO is the EU inter-country supply and use tables published at the
NACE-R2 64-activity detail (A\*64 breakdown), distributed as long-format CSV
flatfiles — one supply file and one use file per reference year.

This vignette walks through the full researcher journey:

1. Obtaining the FIGARO flatfiles and placing them under `$SUBE_FIGARO_DIR`
2. Reading the supply and use flatfiles into a canonical `sube_suts` table with
   `read_figaro()`
3. Building NACE-based mapping tables from the output
4. Running `build_matrices()` to get aggregated supply, use, and final-demand
   matrices
5. Running `compute_sube()` to obtain Leontief multipliers
6. Optionally running `estimate_elasticities()` when sidecar VA/EMP/CO2 data
   is available
7. Executing the gated integration test locally

Every code chunk is `eval = FALSE` so the vignette renders cleanly on CRAN and
in CI without requiring the actual FIGARO flatfiles (roughly 400-500 MB each).

This vignette covers FIGARO-specific import. Once data is in canonical
`sube_suts` format, the downstream pipeline — `build_matrices()`,
`compute_sube()`, and `estimate_elasticities()` — works the same for any SUT
source. Only steps 2–4 (file layout, `read_figaro()`, and NACE mapping) are
FIGARO-specific.

> **Tip.** For a one-call equivalent that chains `read_figaro()` through
> `compute_sube()`, see
> [`run_sube_pipeline(source = "figaro", ...)`](../reference/run_sube_pipeline.html)
> and the [Pipeline Helpers vignette](pipeline-helpers.html).

# 2. Obtaining the data

FIGARO is distributed as two flat-format CSV files per reference year.
The expected layout under `$SUBE_FIGARO_DIR` is:

```
$SUBE_FIGARO_DIR/
  flatfile_eu-ic-supply_25ed_{YEAR}.csv
  flatfile_eu-ic-use_25ed_{YEAR}.csv
```

Typical file size is 400-500 MB per file. Once downloaded and placed in the
directory, set the environment variable before running any gated tests:

```bash
export SUBE_FIGARO_DIR=/path/to/figaro
```

`read_figaro()` and the gated test both read this variable at runtime. When the
variable is unset, the test skips; when it points to a directory without the
expected files, `read_figaro()` stops with an informative message.

# 3. Reading the flatfile

For copy-paste purposes, this section uses the synthetic fixture shipped with
the package. No download is required:

```{r}
dir <- system.file("extdata", "figaro-sample", package = "sube")
sut <- read_figaro(dir, year = 2023)
str(sut)
#> Classes 'sube_suts' and 'data.table':  ...
#>  $ REP : chr  "AT" "BE" ...
#>  $ PAR : chr  "AT" "BE" ...
#>  $ CPA : chr  "A01" "A01" ...
#>  $ VAR : chr  "A01" "A01" ...
#>  $ VALUE: num  ...
#>  $ YEAR : int  2023 2023 ...
```

`read_figaro()` does three things at import time:

- **Primary-input row filtering:** rows where `VAR` is one of `B2A3G`, `D1`,
  `D21X31`, `D29X39`, `OP_RES`, or `OP_NRES` are dropped. These are
  value-added blocks, not product flows, and should not appear in the supply or
  use matrices.
- **`CPA_` prefix stripping:** the raw FIGARO product column carries codes like
  `CPA_A01`. The prefix is removed at import, so the `CPA` column in the output
  contains plain NACE-R2 codes (e.g. `A01`). Downstream code never sees the
  prefix.
- **Final-demand aggregation:** the five FIGARO final-demand codes (`P3_S13`,
  `P3_S14`, `P3_S15`, `P51G`, `P5M`) are summed into a single `VAR = "FU_bas"`
  row per `(REP, PAR, CPA)` so the output is compatible with `build_matrices()`
  without modification.

To use real FIGARO data, replace the `system.file()` path with the env-var
pointer:

```{r}
root <- Sys.getenv("SUBE_FIGARO_DIR")
sut  <- read_figaro(root, year = 2023)
```

# 4. Preparing mapping tables

FIGARO uses NACE-R2 3-digit codes for both products (the `CPA` column, after
prefix stripping) and industries (the `VAR` column for supply flows). Because
NACE product codes and industry codes are equivalent at the 3-digit level,
**one aggregation table serves as both `cpa_map` and `ind_map`** for
`build_matrices()`.

The mapping can be as simple as collapsing all codes to their NACE section
letter (A-U):

```{r}
domestic <- extract_domestic_block(sut)

# Collect all unique codes appearing in product and industry positions
codes <- sort(unique(c(domestic$CPA, setdiff(domestic$VAR, "FU_bas"))))

# Map each 3-digit code to its NACE section letter (first character)
cpa_map <- data.table::data.table(
  CPA    = codes,
  CPAagg = substr(codes, 1, 1)
)

ind_map <- data.table::data.table(
  NACE   = codes,
  INDagg = substr(codes, 1, 1)
)
```

Note that the column name differs between the two maps: `CPA` vs. `NACE`.
`build_matrices()` routes each internally via `.coerce_map()`, which recognises
`NACE` and `NACE_R2` as synonyms for the industry-identifier column. Using
different column names ensures each map routes correctly without ambiguity.

For finer resolution you may use NACE 2-digit codes instead:
`substr(codes, 1, 2)`. The section-level (1-letter) aggregation is simplest
and produces matrices large enough for non-degenerate Leontief inversion across
the full A\*64 country set.

# 5. Building matrices

```{r}
bundle <- build_matrices(domestic, cpa_map, ind_map)

names(bundle)
#> [1] "aggregated"    "final_demand"  "matrices"

# Inspect the supply matrix for Germany
b_de <- bundle$matrices[["DE_2023"]]
dim(b_de$S)
#> [1] 21 21
```

`build_matrices()` returns a named list with:

- `aggregated`: the aggregated long-format supply and use table after applying
  the mapping.
- `final_demand`: the aggregated final-demand totals by `(country, agg_sector)`.
- `matrices`: a named list of per-country-year objects, each containing the
  supply matrix `S`, use matrix `U`, and metadata fields `year`, `country`, and
  `industries`.

# 6. Computing multipliers

`compute_sube()` takes the matrix bundle and an inputs table, inverts
`(I - A)` for each country-year, and returns multipliers and a tidy results
table. When sidecar data (VA, EMP, CO2) is unavailable — the default for
FIGARO — synthesise a gross-output vector from the supply matrix column sums:

```{r}
inputs <- data.table::rbindlist(lapply(names(bundle$matrices), function(nm) {
  b <- bundle$matrices[[nm]]
  data.table::data.table(
    YEAR     = b$year,
    REP      = b$country,
    INDUSTRY = b$industries,
    GO       = as.numeric(colSums(b$S))
  )
}))

result <- compute_sube(bundle, inputs, metrics = "GO")

names(result)
#> [1] "summary"    "model_data" "matrices"

result$summary[REP == "DE" & YEAR == 2023][1:5]
#> Tidy table with one row per (REP, YEAR, INDUSTRY) and multiplier columns.
```

The multipliers in `result$summary` are deterministic given the supply and use
matrices plus the gross-output vector. They are the quantities the gated test
snapshots for reproducibility verification.

# 7. Extending to elasticities

`estimate_elasticities()` fits the SUBE regression models (OLS, pooled, between
estimators) and requires per-industry gross output and at least one of value
added (`VA`), employment (`EMP`), or carbon emissions (`CO2`). FIGARO's supply
and use flatfiles do not include these sidecar variables; researchers must
supply them from an independent source.

If you have per-industry sidecar data in CSV files with columns
`INDUSTRY, GO, VA, EMP, CO2`, point the optional env var at the directory:

```{r}
# Set SUBE_FIGARO_INPUTS_DIR to a directory of per-country CSV files:
#   DE_2023.csv, FR_2023.csv, IT_2023.csv, NL_2023.csv
inputs_dir <- Sys.getenv("SUBE_FIGARO_INPUTS_DIR", unset = "")
if (nzchar(inputs_dir)) {
  # Load and combine sidecar files (researcher-supplied)
  sidecar <- data.table::rbindlist(lapply(
    list.files(inputs_dir, pattern = "\\.csv$", full.names = TRUE),
    data.table::fread
  ))

  # Replace the synthetic GO vector with researcher data
  result_full <- compute_sube(bundle, sidecar, metrics = c("GO", "VA", "EMP"))
  elasticities <- estimate_elasticities(result_full)
}
```

The opt-in path does not have golden-snapshot assertions in the gated test —
only structural invariants (pipeline completes, coefficients non-NA, signs
plausible) because snapshot assertions on researcher-supplied regression inputs
would not be reproducible across environments.

# 8. Running the gated test locally

With `SUBE_FIGARO_DIR` set to the directory containing both flatfiles:

```bash
SUBE_FIGARO_DIR=/path/to/figaro \
  Rscript -e 'devtools::test(filter = "figaro-pipeline")'
```

The first run captures a testthat golden snapshot (stored in
`tests/testthat/_snaps/figaro-pipeline/`) and emits a note
`Adding new snapshot`. Subsequent runs compare against the snapshot and fail on
numeric drift.

Without the env var (or on CRAN/CI), the tests still run but the gated block
auto-skips while the synthetic-fixture contract test always runs:

```bash
Rscript -e 'devtools::test(filter = "figaro-pipeline")'
#> FIG-E2E-02 synthetic contract: PASS
#> FIG-E2E-01 gated real-data: SKIP (SUBE_FIGARO_DIR not set — FIGARO E2E test skipped)
```

The skip message is deterministic: `"SUBE_FIGARO_DIR not set — FIGARO E2E test skipped"`.

# 9. What is NOT covered in v1.2

- **FIGARO SIOT (product-by-product) tables.** Only industry-by-industry SUTs
  are scoped for v1.2. SIOT support is a future-milestone candidate.
- **Auto-download helpers.** Network access and licensing constraints prevent
  bundling a download helper in the package. Researchers obtain the flatfiles
  directly and point `SUBE_FIGARO_DIR` at the download location.
- **Multi-year batch processing.** Running the full pipeline across multiple
  years or a wider country set is facilitated by the upcoming CONV-* convenience
  helpers planned for a later milestone. The workflow above runs one year at a
  time.
