---
title: "Getting Started with sube"
vignette: >
  %\VignetteIndexEntry{Getting Started with sube}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

`sube` packages the core SUBE workflow into a small set of reusable functions.
It is intended both as a practical toolkit for supply-use based econometrics
and as a companion package to the Stehrer, Rueda-Cantuche, Amores, and Zenz
paper on confidence intervals for input-output multipliers.

The central idea is that rectangular supply and use tables can support a more
flexible econometric treatment of multipliers than point-estimate Leontief
analysis alone, especially when combined with panel-style data structures.

## Load sample inputs

```{r}
sut <- sube_example_data("sut_data")
cpa_map <- sube_example_data("cpa_map")
ind_map <- sube_example_data("ind_map")
inputs <- sube_example_data("inputs")
```

## Build matrices

```{r}
domestic <- extract_domestic_block(sut)
bundle <- build_matrices(domestic, cpa_map, ind_map)
names(bundle)
```

## Compute results

```{r}
result <- compute_sube(bundle, inputs)
head(result$summary)
```

## Estimate elasticity models

```{r}
model_data <- sube_example_data("model_data")
models <- estimate_elasticities(model_data, predictor_vars = c("P01", "P02"))
head(models$tidy)
```

## Relation to the companion paper

The package workflow mirrors the broad structure used in the companion paper:

1. start from supply and use data,
2. derive domestic and aggregated matrices,
3. compute Leontief-style benchmark results, and
4. estimate SUBE regressions on prepared modeling tables.

The shipped example data is intentionally tiny, but the workflow is meant to
scale to larger inter-country and multi-year settings of the kind discussed in
the paper.
