---
title: "Getting Started with sube"
vignette: >
  %\VignetteIndexEntry{Getting Started with sube}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

`sube` packages the core SUBE workflow into a small set of reusable functions.
It is intended both as a practical toolkit for supply-use based econometrics
and as a companion package to the Stehrer, Rueda-Cantuche, Amores, and Zenz
paper on confidence intervals for input-output multipliers.

Companion paper:
[Wrapping input-output multipliers in confidence intervals](https://doi.org/10.1186/s40008-024-00331-4),
*Journal of Economic Structures* 13, 17 (2024).

The central idea is that rectangular supply and use tables can support more
than point-estimate Leontief analysis alone. With the right data preparation,
they can also support econometric multiplier work, comparison plots, and a more
explicit treatment of uncertainty across countries, products, and years.

That paper motivation matters for the package design. `sube` does not treat the
econometric layer as a side calculation after Leontief analysis; it treats
Leontief benchmarks, SUBE estimates, and comparison plots as one coherent
research workflow.

This vignette shows the compact public workflow used across the package docs:

1. import or load supply-use data,
2. build domestic matrices,
3. compute Leontief benchmark results,
4. estimate SUBE models, and
5. compare or export both layers from package objects.

All examples below use only the shipped sample data, so the full workflow can
be verified from a clean checkout without external downloads. The package
works with any supply-use data in the canonical long format - WIOD and FIGARO
are two built-in importers, but you can also bring your own source. Readers
who want to skip ahead to input contracts and format details can jump to the
[data-preparation vignette](data-preparation.html).

## Load sample inputs

The shipped example objects map directly to the workflow stages: `sut_data` for supply-use tables, `cpa_map` and `ind_map` for aggregation, `inputs` for the compute layer, and `model_data` for SUBE estimation.

```{r}
sut <- sube_example_data("sut_data")
cpa_map <- sube_example_data("cpa_map")
ind_map <- sube_example_data("ind_map")
inputs <- sube_example_data("inputs")
```

## Build matrices

```{r}
domestic <- extract_domestic_block(sut)
bundle <- build_matrices(domestic, cpa_map, ind_map)
names(bundle)
```

The matrix bundle should expose the aggregated long data, final demand, and the
country-year matrix list used by the compute layer.

```{r}
names(bundle$matrices)
```

## Compute Leontief benchmark results

```{r}
result <- compute_sube(bundle, inputs)
head(result$summary)
```

The returned object contains both tidy outputs and the underlying matrix layer.

```{r}
names(result)
names(result$matrices[[1]])
result$diagnostics
```

With the shipped sample data, the diagnostics table should report `ok`. On real
inputs, this table is the first place to check for singular supply, gross
output, or Leontief branches.

## Estimate SUBE models

The shipped `model_data` is a tiny synthetic fixture, so `estimate_elasticities()`
emits `NaN`/"essentially perfect fit" warnings here. Those are artefacts of the
small sample - real datasets do not trigger them. Warnings are suppressed below
to keep the output readable.

```{r, warning = FALSE}
model_data <- sube_example_data("model_data")
models <- estimate_elasticities(model_data, predictor_vars = c("P01", "P02"))
head(models$tidy)
```

## Compare Leontief and SUBE outputs

```{r}
comparison <- prepare_sube_comparison(
  leontief = result,
  models = models,
  measure = "multiplier",
  variables = c("GO", "VA")
)

head(comparison)
```

The comparison layer is designed to support the same country, product, density, regression, and export paths described in the README and modeling/output vignette.

```{r, eval = FALSE}
country_plots <- plot_paper_comparison(
  comparison,
  kind = "by_country",
  measure = "multiplier",
  variables = "GO"
)

paper_regression <- plot_paper_regression(
  comparison,
  method = "between",
  measure = "multiplier",
  variables = "GO"
)

write_sube("outputs/benchmark_summary", result$summary, format = "csv")
```

## Relation to the companion paper

The package workflow mirrors the broad structure used in the companion paper:

1. start from supply and use data,
2. derive domestic and aggregated matrices,
3. compute Leontief-style benchmark results,
4. estimate SUBE regressions on prepared modeling tables, and
5. compare benchmark and econometric results with common plotting layouts.

The shipped example data is intentionally tiny, but the workflow is meant to
scale to larger inter-country and multi-year settings of the kind discussed in
the paper.
