---
title: "Modeling and Outputs"
vignette: >
  %\VignetteIndexEntry{Modeling and outputs}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(sube)
```

This article covers the tidy result layer, filtering, plots, and export.

It also shows how the package supports the comparison logic used in the
companion paper,
[Wrapping input-output multipliers in confidence intervals](https://doi.org/10.1186/s40008-024-00331-4).
The key idea is to keep Leontief benchmark results and SUBE model outputs in a
shared comparison structure instead of treating them as separate reporting
pipelines.

The examples below use shipped sample data, but the modeling and output layer
works identically for any SUT source once data is in the canonical long format
- whether it came from WIOD, FIGARO, or a custom national supply-use table.
Data preparation and format details are covered in the
[data-preparation vignette](data-preparation.html).

## Compute the sample workflow

```{r}
bundle <- build_matrices(
  extract_domestic_block(sube_example_data("sut_data")),
  sube_example_data("cpa_map"),
  sube_example_data("ind_map")
)
result <- compute_sube(bundle, sube_example_data("inputs"))
head(result$tidy)
```

## Filter and plot

```{r}
filtered <- filter_sube(result$tidy)
plot_sube(filtered, by = "country", kind = "boxplot", measure = "multiplier")
```

## Paper-style comparison plots

```{r}
models <- estimate_elasticities(
  sube_example_data("model_data"),
  predictor_vars = c("P01", "P02")
)

comparison <- prepare_sube_comparison(
  result,
  models,
  measure = "multiplier",
  variables = c("GO", "VA")
)

leontief_long <- extract_leontief_matrices(result, matrix = "L", format = "long")
head(leontief_long)

names(plot_paper_comparison(
  comparison,
  kind = "by_country",
  measure = "multiplier",
  variables = "GO"
))
```

`extract_leontief_matrices()` also supports `format = "list"` and `format = "wide"` when you want matrix-oriented outputs instead of the tidy long table used by most comparison helpers.

These helpers are designed for the same kinds of questions emphasized in the
paper: where Leontief and econometric estimates line up, where they diverge,
and how those differences vary across countries and products.

## Export

```{r, eval = FALSE}
write_sube("outputs/final_results", filtered, format = "csv")
write_sube(
  "outputs/comparison_bundle",
  list(filtered = filtered, summary = result$summary),
  format = "csv"
)
```

A single table writes to one normalized file path. A named list writes one file per entry into the target directory, which is useful for comparison-ready bundles.

## Model estimation tables

```{r}
models <- estimate_elasticities(
  sube_example_data("model_data"),
  predictor_vars = c("P01", "P02")
)
head(models$pooled)
```

The comparison layer can also be used for regression-style and interval-range
plots when the modeling results contain enough information. These helpers return named lists of `ggplot` objects keyed by the available model or variable selections in the supplied data:

```{r}
paper_regression <- plot_paper_regression(
  comparison,
  method = "between",
  measure = "multiplier",
  variables = "GO"
)

interval_ranges <- plot_paper_interval_ranges(
  models,
  by = "country",
  variables = "GO"
)
```
