---
title: "Reproducing the SUBE Paper"
vignette: >
  %\VignetteIndexEntry{Reproducing the SUBE Paper}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  eval = FALSE
)
library(sube)
```

# 1. What this vignette replicates

The 2018 SUBE paper computes Leontief multipliers, elasticities, and
regression coefficients from WIOD international supply-use tables. This
vignette walks through the end-to-end reproduction of the paper's raw
supply, use, and net-supply matrices using only the functions exported
by `sube`. All code chunks are shown `eval = FALSE` so the vignette
builds cleanly on CRAN and in CI without requiring the ~4 GB WIOD data
archive.

The gated test at `tests/testthat/test-replication.R` asserts bit-level
numerical equivalence for a representative subset (AUS, DEU, USA, JPN
for 2005). Running it locally is covered in section 9.

# 2. Obtaining the WIOD data

Download the WIOD international supply-use tables and companion files
from Eurostat / the WIOD website. The expected layout under
`$SUBE_WIOD_DIR`:

```
$SUBE_WIOD_DIR/
  International SUTs domestic/
    Int_SUTs_domestic_SUP_2005_May18.csv
    Int_SUTs_domestic_USE_2005_May18.csv
    ...
  Correspondences/
    CorrespondenceCPA56.dta
    CorrespondenceInd56.dta
  GOVAcur/      GO_{country}_{year}.dta
  EMP/          EMP_{country}_{year}.dta
  CO2/          CO2_{country}_{year}.dta
  Regression/data/   {country}_{year}.csv  (legacy ground-truth W matrices)
```

Set the environment variable before running the gated test:

```bash
export SUBE_WIOD_DIR=/path/to/wiod
```

# 3. Importing the domestic block

```{r}
sut <- import_suts(file.path(Sys.getenv("SUBE_WIOD_DIR"),
                             "International SUTs domestic"))
domestic <- extract_domestic_block(sut)
domestic[REP == "AUS" & YEAR == 2005][1:3]
#> Returns a sube_domestic_suts long table with REP == PAR rows only.
```

# 4. Aggregation via correspondence tables

```{r}
root <- Sys.getenv("SUBE_WIOD_DIR")
cpa_map <- data.table::data.table(haven::read_dta(
  file.path(root, "Correspondences", "CorrespondenceCPA56.dta")))
ind_map <- data.table::data.table(haven::read_dta(
  file.path(root, "Correspondences", "CorrespondenceInd56.dta")))
data.table::setnames(cpa_map, "CPAagg", "CPA_AGG")
data.table::setnames(ind_map, "Indagg", "IND_AGG")
```

The CPA56 and Ind56 maps project the raw 56-sector NACE codes onto the
22-sector aggregation used throughout the paper.

# 5. Computing Leontief multipliers and elasticities

```{r}
results <- compute_sube(domestic, cpa_map, ind_map)
results$tidy[COUNTRY == "AUS" & YEAR == 2005 & measure == "multiplier"][1:5]
```

`compute_sube()` builds the aggregated S and U matrices, inverts
`(I - A)`, and emits tidy multiplier and elasticity tables.

# 6. Applying the paper's outlier treatment

```{r}
models <- estimate_elasticities(results)
comp <- prepare_sube_comparison(results, models)
comp_filtered <- filter_paper_outliers(comp)
#> Applies the six exclusion layers from 08_outlier_treatment.R:89-181 --
#> see ?filter_paper_outliers for the full rule list.
```

Pass `apply_bounds = FALSE` to keep multiplier outliers in the output,
or pass `variables = c("GO", "VA", "EMP")` to skip the CO2-specific
rules when CO2 data is unavailable.

# 7. Running the regressions

```{r}
models <- estimate_elasticities(results)
models$OLS[term == "P01" & variable == "GO"]
#> The SUBE paper's OLS, pooled, and between estimators are all produced
#> in one call. See ?estimate_elasticities for the full return shape.
```

# 8. Comparing with the legacy paper output

Running the full pipeline on the 2018 WIOD data yields:

```
#> Raw supply and use matrices: exact match (bit-identical)
#> Net-supply model matrix W = t(SUP_agg - USE_agg): exact match
#> Leontief multipliers after outlier treatment: ~2.7% mean abs. diff.
#> OLS coefficients (significant terms): match to 4+ decimal places
```

The matrix-level differences are zero; the multiplier- and
coefficient-level differences are methodological (averaging order,
filter interaction) and are documented in full in
`inst/scripts/replicate_paper.R`.

# 9. Running the gated test locally

With `SUBE_WIOD_DIR` pointing at a populated WIOD tree:

```bash
SUBE_WIOD_DIR=/path/to/wiod Rscript -e 'devtools::test(filter = "replication")'
```

Without the env var (or on CRAN / CI), the test auto-skips:

```bash
Rscript -e 'devtools::test(filter = "replication")'
#> SKIP (SUBE_WIOD_DIR not set - paper replication test skipped)
```

# Beyond this vignette

- **IHS regression variants.** `archive/legacy-scripts/06_SUBE_ihs*.R`,
  `06_SUBE_lin-ihs*.R`, `06_SUBE_ihs-lin*.R` fit inverse-hyperbolic-sine
  transformed variants. Not part of `sube` v1.1; use the legacy scripts
  directly if needed.
- **Paper figures.** `plot_paper_comparison()`, `plot_paper_regression()`,
  and `plot_paper_interval_ranges()` reproduce the published figures.
  Each ships with its own man page.
- **Full comparison runbook.** `inst/scripts/replicate_paper.R` is the
  reference runbook; this vignette is the narrated subset.
