---
title: "Pipeline Helpers — One Call from Path to Multipliers"
vignette: >
  %\VignetteIndexEntry{Pipeline Helpers — One Call from Path to Multipliers}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  eval = TRUE
)
library(sube)
library(data.table)
```

# 1. When to reach for the convenience helpers

The four-step SUBE chain — `import_suts()` / `read_figaro()` →
`extract_domestic_block()` → `build_matrices()` → `compute_sube()` — gives you
full control over every stage. If you already have the intermediate objects in
hand, call the four steps directly; it is two extra lines.

Reach for the convenience helpers in two cases:

1. You have a SUT file (or directory) on disk and want multipliers back in one
   call: use `run_sube_pipeline()`.
2. You have a pre-imported `sube_suts` table and want to sweep across many
   country-year slices, collecting tidy merged outputs: use `batch_sube()`.

Both helpers surface the same structured diagnostics contract so silent
data-quality issues (dropped rows, misaligned inputs, singular branches) never
disappear from the console.

# 2. `run_sube_pipeline()` on sample data

The package ships a tiny WIOD-format sample under
`system.file("extdata", "sample", package = "sube")`. Feed its `sut_data.csv`
straight into `run_sube_pipeline()`:

```{r}
sut_path <- system.file("extdata", "sample", "sut_data.csv", package = "sube")

result <- run_sube_pipeline(
  path    = sut_path,
  cpa_map = sube_example_data("cpa_map"),
  ind_map = sube_example_data("ind_map"),
  inputs  = sube_example_data("inputs"),
  source  = "wiod"
)

class(result)
names(result)
```

# 3. Inspecting `$results`, `$diagnostics`, and `$call`

The compute output lives under `$results`:

```{r}
result$results$summary
```

The unified diagnostics table has a 6-column schema — `country`, `year`,
`stage`, `status`, `message`, `n_rows`:

```{r}
result$diagnostics
```

When every status is `"ok"` no warning is emitted. Provenance about the call
itself lives under `$call`:

```{r}
result$call[c("source", "path", "n_countries", "n_years", "estimate",
              "package_version")]
```

# 4. Switching to the FIGARO path

The same helper imports FIGARO flatfiles when `source = "figaro"`. Because
FIGARO data is ~400 MB per flatfile, the block below is illustrative
(`eval = FALSE`):

```{r eval = FALSE}
figaro_dir <- Sys.getenv("SUBE_FIGARO_DIR")   # user-provided
run_sube_pipeline(
  path    = figaro_dir,
  cpa_map = my_cpa_map,     # user-supplied correspondence
  ind_map = my_ind_map,
  inputs  = my_inputs,      # industry-level GO/VA/EMP/CO2
  source  = "figaro",
  year    = 2023L
)
```

See `vignette("figaro-workflow")` for the full FIGARO narrative including the
env-var gate (`SUBE_FIGARO_DIR`) and map-table construction.

# 5. `batch_sube()` across groups

The sample data has one country (AAA) and one year (2020). To illustrate
multi-group batching, duplicate the sample into a second year:

```{r}
sut <- sube_example_data("sut_data")
sut2 <- data.table::copy(sut); sut2[, YEAR := 2021L]
sut_multi <- rbind(sut, sut2)
class(sut_multi) <- c("sube_suts", class(sut_multi))

inp <- sube_example_data("inputs")
inp2 <- data.table::copy(inp); inp2[, YEAR := 2021L]
inp_multi <- rbind(inp, inp2)

batch <- batch_sube(
  sut_data = sut_multi,
  cpa_map  = sube_example_data("cpa_map"),
  ind_map  = sube_example_data("ind_map"),
  inputs   = inp_multi
)

names(batch$results)
batch$summary
batch$diagnostics
```

Per-group results are preserved under `batch$results`; merged tables under
`$summary`, `$tidy`, and `$diagnostics` (with a `group_key` column naming
each batch key) are ready for downstream analysis.

# 6. Turning on `estimate = TRUE`

When you also want elasticities from `estimate_elasticities()`, pass
`estimate = TRUE`. The sample data is tiny so the regression may return
`NULL`; on production-sized inputs a `sube_models` object is attached:

```{r}
result_est <- run_sube_pipeline(
  path     = sut_path,
  cpa_map  = sube_example_data("cpa_map"),
  ind_map  = sube_example_data("ind_map"),
  inputs   = sube_example_data("inputs"),
  source   = "wiod",
  estimate = TRUE
)
class(result_est$models)   # "sube_models" when model_data is non-empty, else NULL
```

# 7. Reading the diagnostic warnings

The `$diagnostics` table categorises silent issues into four statuses:

- **`coerced_na`** (`stage = "import"`) — rows whose `VALUE` became `NA`
  during `as.numeric()` coercion at import. Pipeline-level aggregate; count
  lives in `n_rows`.
- **`skipped_alignment`** (`stage = "build"`) — country-years present in
  the SUT input but dropped by `build_matrices()` because the CPA or
  industry correspondence left no aligned rows.
- **`inputs_misaligned`** (`stage = "build"`) — country-years present in
  both the SUT data and the `inputs` table but absent from
  `build_matrices()`'s `$model_data`. Typical cause: industry codes in
  `inputs` that do not match the matrix industries.
- **`singular_supply` / `singular_go` / `singular_leontief`**
  (`stage = "compute"`) — `compute_sube()` could not invert the relevant
  matrix for this country-year. Pass-through from the compute stage's
  existing diagnostics.

When any row has `status != "ok"`, a single `warning()` summarises counts
per category. Query the structured table for programmatic follow-up; read
the warning for a fast interactive heads-up.
